---
title: "MAS 637 - Applied Regression Analysis 1 Final Project"
author: "Qian Dong, Xiaoxi Yuan, Eduardo Santiago, & Philip Bachas-Daunert"
date: "Due Date: 2023-10-06"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    keep_tex: true
    fig_caption: true
    latex_engine: xelatex
header-includes:
  - "\\usepackage{sectsty}"
  - "\\allsectionsfont{\\centering}"
  - "\\geometry{legalpaper}" # Setting the geometry here
---

```{r setup, include=FALSE}
#setting up the knitr for other chunks
knitr::opts_chunk$set(
    eval = TRUE,
    message = FALSE,
    warning = FALSE,
    comment = NA,
    echo = FALSE   # This line ensures code is not shown in the output
)

#loading libraries in this chunk and setting things up.
library(tidyr)
library(ggplot2)
library(dplyr)

#Quick cleaning of the dataset.
d = read.csv('Walmart.csv')
d$Date = as.Date(d$Date, '%m/%d/%y')
```

\newpage

# Abstract

Walmart as global retail leader has at its disposal a vast array of data
reflecting its day-to-day operations which provide key insights into how
different aspects of its business affect each other. Through this
project, our group aims to make use of Walmart's extensive operational
data to help determine what are the factors at play which affect
Walmart's sales, we will also substantiate the holiday sales claims made
within the problem statement by hypothesis testing, and search for
significant differences in sales across the different store types. The
primary objectives of our analysis will be met by conducting different
analysis techniques learned in class ranging from: testing for
significance amongst different variables, interpreting the impact of
categorical vs numerical variables of a data set, making use of
different plots to interpret our data in a visual point of reference,
conduct hypothesis testing on claims made in the initial problem
statement, and much more.

# Introduction

With a market capitalization of over \$430 billion, and over 4,600
stores in just the United States alone, Walmart by the very nature of
its business is a treasure trove of operational data which lends itself
to a wide array of statistical analysis to draw key insights into its
business and its strategic decision making. The motivation behind
selecting the Walmart case to conduct our final project stems from the
valuable insight our group will be able to derive from analyzing the
underlying factors affecting the sales performance of a company like
Walmart, who is a leader in its industry, and in turn we will be well
equipped when entering the labor market to make similar analysis for
other companies' key performance metrics, revenue drivers, growth
opportunities and much more. Through this project, our group will tackle
different questions posed by the problem statement, first we will
develop a parsimonious regression model to be able to interpret the
variables at play which drive Walmart's sales. Second, we will address
the holiday sales claim made that during the holiday season, sales are
expected to be 10% more than when compared to non-holiday sales. More
specifically, we will implement hypothesis testing techniques learned in
class to validate this assertion and generate a confidence interval for
the holiday sales coefficient ("IsHoliday"). Lastly, we will ascertain
if in fact sales are significantly different across different store
types and answer this claim with a statistical test. As we initiate our
statistical analysis on Walmart's sales, we hope to successfully employ
different techniques learned in class to draw actionable insights on
this problem set and in turn allow us to sharpen our analytical toolbox
which has become so coveted in today's vastly changing business
landscape.

# Analysis

We obtained the sales data of 45 Walmart stores from 2010 to 2012. From
this data set, we can see the factors that may affect the sales revenue
of Walmart stores: weather temperature, oil price, CPI, unemployment
rate, whether it is a holiday, sales quantity, and store type. The
average revenue of the store is \$1,046,965, with a standard deviation
of \$564,366. After the initial descriptive analysis, we can calculate
the correlation matrix to assess the linear relationships between the
variables The correlation matrix helps us understand which independent
variables are significantly correlated with the dependent variables and
with each other Scatter plots are a useful tool for visualizing the
relationships between the dependent and independent variables We also
create scatter plots for pairs of variables to examine how they relate
to each other Then We processed with linear regression modeling using
the lm() function to build and interpret regression models.

# Modeling

The model is defined as follows:
`lm(Sales ~ Temperature + Fuel_Price + CPI + Unemployment + IsHoliday + Size + StoreType, data = d)`.

\newpage

# Bullet 1: Interpret the coefficients in the final model.

```{r}
#This set of code is for creating regression model and summary
reg = lm(Sales ~ Temperature + Fuel_Price + CPI + Unemployment + 
           IsHoliday + Size + StoreType, data = d)
summary(reg)
```

The following variables are significant:

-   **CPI**

    -   Other things remain unchanged, each additional change in CPI is
        associated with a decrease of approximately \$1,390 in sales.

-   **Fuel_Price**

    -   Other things remain unchanged, each additional change in fuel
        price is associated with a decrease of approximately \$29,960 in
        sales.

-   **Size**

    -   Other things remain unchanged, each additional change in store
        size is associated with an increase of approximately \$7,929
        sales.

-   **StoreTypeB**

    -   Other things remain unchanged, sales will increase \$48,570 by
        choosing TypeB store.

-   **StoreTypeC**

    -   Other things remain unchanged, sales will increase \$195,000 by
        choosing TypeC store.

-   **Temperature**

    -   Other things remain unchanged, each additional change in
        temperature is associated with an increase of approximately
        \$1,293 in sales.

-   **Unemployment**

    -   Other things remain unchanged, each additional change in
        unemployment is associated with an increase of approximately
        \$25,100 in sales. \newpage

# Bullet 2: Our boss claims holiday sales will be 10% more than on a regular day. Is this claim correct? Conduct a hypothesis test to test the claim. Hint: you may want to generate a confidence interval for the holiday sales coefficient (isholiday).

$H_0: S_{\text{holiday}} = 1.10 \times S_{\text{regular}}$

```{r}
reg2 = lm(Sales ~ IsHoliday, data = d)
summary(reg2)

cf1 = coef(reg2)

intercept1 = cf1[1]
slope1 = cf1[2]
```

$y = 1044714 + 32184 \times x_{\text{IsHoliday}}$

```{r}
conf = confint(reg2)

intercept2 = conf[1]
slope2 = conf[2]
intercept3 = conf[3]
slope3 = conf[4]

sales_min_holiday = intercept2 + slope2
sales_max_holiday = intercept3 + slope3

percent_min = sales_min_holiday/intercept1 - 1
percent_max = sales_max_holiday/intercept1 - 1
```

```{r}
cat('The min percent is: ', sprintf("%.3f%%", 100*percent_min))
cat('The max percent is: ', sprintf("%.3f%%", 100*percent_max))
```

Since the highest percentage of increase in sales in terms of whether it
was a holiday or not was 9.626%, we failed to accept Hypothesis $H_0$
which claims that the sales in holiday will be 10% less than a regular
day. \newpage

# Bullet 3: It is believed that sales are significantly different across the different store types. Can you refute this claim using a statistical test?

```{r}
dummy_stores = factor(d$StoreType)

reg3 = lm(Sales ~ dummy_stores, data = d)
summary(reg3)
```

We assume different store types will influence significantly on sales.
We made a regression model to test it. The results showed that all types
of store A, B, C have a $p\text{-value} < 0.05$ , which means we
couldn't refute the claim. \newpage

# Diagnostics

```{r}
# The histogram of the residuals below shows that the residuals are approximately normally distributed and skewed to the right.

# Create a data frame to store the residuals
residuals_data <- data.frame(residuals = reg$residuals)

# Create a ggplot histogram of residuals
p <- ggplot(residuals_data, aes(x = residuals)) +
  geom_histogram(binwidth = 100000, fill = "#0071CE", color = "black", alpha = 0.7) + 
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency") +
  theme_minimal() +
  geom_vline(xintercept = mean(residuals_data$residuals), linetype = "dashed", color = '#FFC220', size = 1) + 
  annotate("text", x = mean(residuals_data$residuals) + 350000, y=Inf, vjust = 2, label = paste("Mean =", round(mean(residuals_data$residuals), 2)), color = "black") +
  theme(text = element_text(size = 15),
        plot.title = element_text(hjust = 0.5, face = "bold"))

# Fun fact, the hexcode above are actually Walmart's blue and yellow.

# Display the plot
print(p)

```

\newpage

# Conclusion

After concluding our analysis of the Walmart Sales data, our team was
able to answer the following questions posed by the problem set: 

1.  Through analyzing the coefficients of the model, we were able to
    determine that the following variables were statistically
    significant due to them having a p-value below 0.05 (CPI,
    Fuel_Price, Size, the dummy variables introduced: StoreTypeB &
    StoreTypeC, Temperature, and Unemployment). The model also had an
    Adjusted R-squared of 0.6719, which indicated a relatively good fit
    of our regression model to the data provided, as 67.23% of the
    variability in the response variable "Sales" is explained by the
    predictor variables in the model. The overall model had a p-value of
    less than 2.2e-16 indicating that there was a very strong level of
    statistical significance amongst the variables analyzed.

2.  The 2^nd^ bullet in the Walmart Sales problem asked us to determine
    whether the claim that holiday sales will be 10% more than
    non-holiday sales. This prompted us to set the null hypothesis:
    Sales(holiday) = 1.10 x Sales(non-holiday) and the alternate
    hypothesis would be refuting this claim. Our team was able to
    predict Sales as a function of the variable "IsHoliday" to arrive at
    the estimates for the intercept and the "IsHoliday" variable. After
    obtaining these data points, we were able to construct a simple
    regression equation which helped us determine that the highest
    percentage increase in sales in terms of whether it was a holiday or
    not was 9.626%, thus we failed to accept the null hypothesis which
    claims that the sales in holiday will be 10% more than a regular
    day.

3.  The last bullet in this problem prompted us to analyze if sales are
    significantly different across the different store types per the
    data in the Walmart.csv file. Again, our team was able make use of
    the techniques learned in class and construct 2 dummy variables
    (dummy_storesB & dummyStoresC) and run a regression model to
    determine the influence of store type on sales. The results showed
    that all types of stores A, B, C have a p-value below 0.05, which
    meant that we couldn't refute the claim that sales are significantly
    influenced by the different store types.
