---
title: "Handout on Simple Linear Regression 2"
author: "Philip Bachas-Daunert"
date: "`r Sys.Date()`"
output: pdf_document
---

# Questions

## 1. A sample of 10 billionaires is selected, and the person’s age and net worth (in billions) are compared. The data are given here: X (age): 56, 39, 42, 60, 84, 37, 68, 66, 73, 55 | Y (net worth in Billions USD): 18, 14, 12, 14, 11, 10, 10, 7, 7, 5

### 1. (a & b) Find $\widehat{y}$ for each x & find a residual for each x.

```{r}
age = c(56, 39, 42, 60, 84, 37, 68, 66, 73, 55)
netB = c(18, 14, 12, 14, 11, 10, 10, 7, 7, 5)

df = data.frame(age , netB)

model = lm(netB~age, data = df)

data.frame(df, y_hat = fitted(model), e = residuals(model))

```

### 1. (c) Calculate the sum of squares total (SST).

```{r}

SST = sum()

```

### 1. (d) Calculate the sum of squares regression (SSR).

```{r}



```

### 1. (e) Calculate the sum of squares due to error (SSE).

```{r}



```

### 1. (f) Calculate the coefficient of determination ($R^{2}$).

```{r}



```

### 1. (g) Calculate the mean square error (MSE).

```{r}



```

### 1. (h) Calculate the root mean square error (RMSE).

```{r}



```

### 1. (i) Calculate the standard error of the slope estimate ($b_{1}$).

```{r}



```

### 1. (j) Calculate the t-value for the slope estimate ($b_{1}$).

```{r}



```

### 1. (k) Calculate the standard error of the intercept estimate ($b_{0}$).

```{r}



```

### 1. (l) Calculate the t-value for the intercept estimate ($b_{0}$).

```{r}



```

### 1. (m) Calculate the t-value for the intercept estimate ($b_{0}$).

```{r}



```

\newpage

## 2. The data on price ($) and the overall score for six stereo headphones tested by *Consumer Reports* were as follows (*Consumer Reports* website, March 5, 2012): Brand: Bose , Skullcandy , Koss , Phillips/O’Neill,  Denon , JVC | Price (USD): 180 , 150 , 95 , 70 , 70 , 35 | Score: 76 , 71 , 61 , 56 , 40 , 26

### 2. (a) Find the linear regression equation that predicts the Score based on Price. State the parameter estimates to 2 places past the decimal.

```{r}



```

### 2. (b) Calculate the SST, SSR, and SSE.


```{r}



```

### 2. (c) Compute the coefficient of determination ($R^{2}$).

```{r}



```

### 2. (d) Compute the root mean square error (RMSE).

```{r}



```

### 2. (e) Calculate the standard error of the slope estimate ($b_{1}$).

```{r}



```

### 2. (f) Perform a t-test.

```{r}



```

### 2. (g) Does the t test indicate a significant relationship between price and the overall score? What is your conclusion?

*type answer here*

### 2. (g) Test for a significant relationship using the F-test. What is your conclusion?

```{r}



```

*type answer here*